% Place acknowledgments below.

This work would not have been possible without the help and support of a wonderful network of people, many of whom have become good friends over the past 8 years.

The work in this dissertation was done in collaboration with my advisors Judith Kroll and Giuli Dussias, as well as with my colleagues Grant Berry and Christian Navarro-Torres. Together, we acknowledge financial support from NSF OISE 0968369 PIRE Grant to the Center for Language Science, NSF Grant BCS-0821924 to Paola Dussias, NIH Grant HD50629 to Paola Dussias, NSF Grant BCS-1331709 to Jason Gullifer, Liberal Arts College Dissertation Support Grant to Jason Gullifer, NSF Grant BCS-0955090 to Judith Kroll, and NIH Grant HD053146 to Judith Kroll.

I would like to thank my research support: Joe Bauman, Mar\'{i}a Luisa Gomez-Marco, Vanessa Murray, Manuel Pulido, Jeremy Slocum, Gabriela Terrazas, and Maya Wade. I would also like to thank Matt Carlson for providing statistical help. I especially thank Anne McPherson for all of her hard work on many of my projects. She did an amazing job keeping everything organized and recruiting and testing so many of the participants.

I would also like to thank my colleagues from other universities who helped collect data for this project: Teresa Bajo, Laura Birke Hansen, and Giulia Togato at La Universidad de Granada as well as Ana Schwartz and Justin Lauro at the University of Texas El Paso. Also, another thanks to Teresa Bajo who was my faculty host during my PIRE trip to Granada and to Ana Schwartz for hosting me three times to collect data over the course of my graduate studies. I hope to visit Granada and El Paso again soon!

I owe much to my lab mates, associates, and friends and thank them for their almost daily support and guidance: Colleen Balukas, Annie Beatty, Cari Bogulski, Kinsey Bice, Amelia Dietrich, Melinda Fricke, Kaitlyn Litcofsky, Fengyang Ma, Bobby Marker, Mari-Cruz Mart\'{i}n, Rhonda McClain, Christian Navarro-Torres, Ji Sook Park, Lauren Perrotti, Tim Poepsel, Nick Ray, Eleonora Rossi, Patricia Rom\'{a}n, Caitlin Ting, Alvaro Villegas, Ben Zinszer, and Megan Zirnstein. I would like to thank Sharon Elder for helping me navigate Penn State bureaucracy and for always being willing to talk about anything.

I am greatly indebted to Grant Berry, Anne Beatty, Christian Navarro-Torres for help in preparing an enormous amount of stimuli and for the many discussions about theoretical aspects of the design. I would also like to thank Mari-Cruz Mart\'{i}n for providing me with some of her materials for this project.

I owe much to my current and former mentors. Judith Kroll and Giuli Dussias, you have both taught me so much about research, writing, and academia (the good and bad). I think it's safe to say that you both are among the most dedicated people that I have ever met, to your students, to your research, and to the Center for Language Science community. I would not have reached this point without your continued support. I have very much enjoyed watching the CLS grow over my time here, and I am glad to have played a role in its development from the ground up (literally). I would also like to thank my undergraduate advisors, Lyn Frazier and Chuck Clifton, at the University of Massachusetts, Amherst for introducing me to the world of psycholinguistics.

Finally, I thank my family and extended family. Mom and Dad, you have both always impressed upon me and Mark the value of learning and education. Whether it was helping with schoolwork and complicated class projects, going to bat for me at school, or ``reminding'' me to fill out FAFSA forms and college applications, you were both always there in your own ways. I remember Dad telling me that he had the feeling since I was very young that I would be involved with school and learning for a long time. At the time, this seemed absurd; I always did well in school, but I didn't particularly enjoy being there. In the end, I don't think any of us realized how right he was.  Mark, my favorite brother, thanks for all your support over the years. Ashley, you've always believed in me even in times where I questioned whether I believed in myself.  It has felt like a long three years living apart that's been made tolerable in no small part by your many visits to State College (though who can blame you when you compare State College to Austin).  I can't wait to start our next chapter together in Montreal! I also thank the DaSilvas and Legares for their support, you all have become family over the years. 
